---
title: "Q-Package: Installation and Usage"
author: "Pradap Konda"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Q-Package: User Manual}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

## Introduction

This document provides instructions to install QPackage in R and how it can be used perform different steps in EMS workflow 

## Installation

QPackage is implemented in R language. We assume that R is already installed in the system, if not R can be installed from [R site](http://cran.r-project.org) . Installing QPackage first requires setting some environment variables and then the package can be installed from GitHub or using a tar ball.

### Environment setup

In order to install QPackage we need to set two environment variables: (1) JAVA_HOME, and (2) CLASS_PATH. JAVA_HOME must be set to the path where JDK is installed. If JDK is not installed, it can be installed from [java site](http://www.oracle.com/technetwork/java/javase/downloads/jdk8-downloads-2133151.html).
For example, in unix systems JAVA_HOME and CLASS_PATH can be set as follows
```
bash$ export JAVA_HOME = ...
bash$ export CLASS_PATH = ...

```
and then execute the following command
```
bash$ R CMD javareconf -e
```

**Note:** Installing a R package in CSL machines (using R at ```/usr/bin/R```) require root permissions, which we donot have. We need to install R packages in a custom directory where there is read/write access. For example, let ```/scratch/pradap/r-libs``` directory be used to install R packages. 

Execute the following command to create the directory

```
bash$ mkdir /scratch/pradap/r-libs
```


### Installing from GitHub

To install QPackage from GitHub, first install ```devtools``` package. It can be installed using the following command in R prompt

```
> install.packages("devtools", lib="/scratch/pradap/r-libs")

```
Once the installation goes through, execute the following command to use the package

```
> library("devtools", lib.loc="/scratch/pradap/r-libs")

```
Now the QPackage can be installed using the following command

```
> with_libpaths(new = "/scratch/pradap/r-libs", install_github("kvpradap/Q-RPackages/QPackage"))

```
Once the installation goes through, execute the following command to use the package
```
> library("QPackage", lib.loc="/scratch/pradap/r-libs")

```
### Installing from Tarball

To install QPackage from tarball, first download the package from [here](http://pages.cs.wisc.edu/~pradap). Lets say that the package is downloaded at ```/scratch/pradap/downloads/QPackage_1.0.tar.gz```. 
Now to install the package, execute the following command

```
> install.packages("/scratch/pradap/downloads/QPackage_1.0.tar.gz", repos = NULL, sourece = TRUE)

```

Once the installation goes through, execute the following command to use the package
```
> library("QPackage", lib.loc="/scratch/pradap/r-libs")

```
**Note**: Instead of including ```lib.loc``` in ```library``` everytime in R session, we can do the following
```
> .libPaths("/scratch/pradap/r-libs")

```
and then execute the following 

```{r}
library(QPackage)
```

## EMS workflow
An end-to-end workflow in EMS  possibly contain the following steps

1. Import data, check constraints export data
2. Browse/understand/clean/transform data
3. Blocking
4. Evaluate/Debug blocking result
5. Matching/Clustering
6. Accuracy estimation
7. Debug
8. Visualization of result and user interaction

The goal of QPackage is to support all the above steps. As a first step a subset of above steps are supported and they are described in the following sections.

## Main classes
There are two main classes in QPackage.

* qtable
* tuple

### qtable

It is used to represent input tables such as walmart product table. It is inherited from R ```data.frame```. It contains an attribute called ```key``` that stores a list of unique identifiers in the table. ```key``` attribute is similar to primary key attribute in SQL tables.  

### tuple

It is used to represent an entity in a table such as product in walmart product table. It is represented as ```named list``` in R. 
In QPackage a tuple can be created as follows

```{r}
t <- tuple("title" = "This is a title", "brand" = "Product brand")

# print the class
print(class(t))

# print contents of t
# title
print(t$title)
# brand
print(t$brand)
```

### Import data
#### Check id

#### Set id

#### Pre-loaded datasets
### Export data

### View data

### Blocking

### Sample data

### Label data

### Create features

### Model selection

### Train model

### Predict labels

 
Note the various macros within the `vignette` setion of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
