---
title: "Q-Package: Installation and Usage"
author: "Pradap Konda"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Q-Package: User Manual}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

## Introduction

This document provides instructions to install QPackage in R and its usage to perform different steps in EMS workflow. 

## Convention
This document uses shell commands and R code blocks for illustration.
The shell commands are shown with an explicit ```bash``` prompt, for example

```
bash$ mkdir /scratch/pradap/r-libs
```

The R code block in this document intermingle R command and its output, so that a user can copy and paste the R command and check its output. 

For example, R command and its output will look like

```{r}
print("Hello World")

```

We can observe that the output starts with ```## ```. 

In some cases, the output is not shown as they were quite large or they open up GUI, such commands are noted in comment as in the following example.

```
View(iris) # output is suppressed
```

## Installation

QPackage is implemented in R. We assume that R is already installed in the system, if not R can be installed from [R site](http://cran.r-project.org) . Installing QPackage has two important steps:(1) Setting some environment variables, and (2) Installing package from GitHub or tarball.

### Environment setup

To install QPackage we need to set two environment variables: (1) JAVA_HOME, and (2) CLASS_PATH. 
JAVA_HOME must be set to the path where JDK is installed. If JDK is not installed, it can be installed from [java site](http://www.oracle.com/technetwork/java/javase/downloads/jdk8-downloads-2133151.html).
CLASS_PATH must include current directory (```.```) in its path.

For example, in unix systems JAVA_HOME and CLASS_PATH can be set from command prompt as follows
```
bash$ export JAVA_HOME=/scratch/pradap/local/share/jdk1.7.0_25
bash$ export CLASS_PATH=.:${CLASS_PATH}

```
After exporting the variables, execute the following command from command prompt

```
bash$ R CMD javareconf -e
```

**Note:** Installing a R package in CSL machines (using R at ```/usr/bin/R```) require root permissions, which we donot have. We need to install R packages in a custom directory where we have read/write access. For example, let ```/scratch/pradap/r-libs``` be the directory that we decide to use for installing R packages and the directory can be created using the following command. 

```
bash$ mkdir /scratch/pradap/r-libs

```


### Installing from GitHub

Installing package from GitHub has two steps: (1) Installing ```devtools``` package, and (2) Use ```install_github``` command to install QPackage from GitHub

```devtools``` package can be installed using the following command in R


```
# set the custom directory to R's search path
.libPaths(c("/scratch/pradap/r-libs/", .libPaths())) 

# install devtools package
install.packages("devtools") # output is suppressed

```
Once the installation goes through, execute the following command to use the package

```
library("devtools") # output is suppressed

```
Now the QPackage can be installed using the following command

```
install_github("kvpradap/Q-RPackage/QPackage") # output is suppressed

```
Once the installation goes through, execute the following command to use the package
```{r}
library("QPackage") 

```
### Installing from Tarball

Installing package from Tar ball has two steps: (1) Downloading the tarball, and (2) Use tarball to install the package

The latest version of QPackage (as tarball) can be downloaded from  [here](http://pages.cs.wisc.edu/~pradap/qpackage/QPackage_1.0.tar.gz). Lets say that the package is downloaded at ```/scratch/pradap/downloads/QPackage_1.0.tar.gz```. 

To install the package, execute the following command

```
# set the custom directory to R's search path
.libPaths(c("/scratch/pradap/r-libs/", .libPaths())) 

# install rJava package
install.packages("rJava") # output is suppressed

install.packages("/scratch/pradap/downloads/QPackage_1.0.tar.gz", 
          repos = NULL, source = TRUE) # output is suppressed

```

Once the installation goes through, execute the following command to use the package

```
library("QPackage") # output is suppressed

```
**Note**: Installation from GitHub is recommended as it automatically installs all the dependency packages.

## EMS workflow
An end-to-end workflow in EMS  possibly contain the following steps

1. Import data, check constraints export data
2. Browse/understand/clean/transform data
3. Blocking
4. Evaluate/Debug blocking result
5. Matching/Clustering
6. Accuracy estimation
7. Debug
8. Visualization of result and user interaction

The goal of QPackage is to support all the above steps. As a first step a subset of above steps are supported and they are described in the following sections.

## Main classes
There are two main classes in QPackage.

* qtable
* tuple

### qtable

It is used to represent input tables such as walmart product dataset. It is inherited from R's ```data.frame```. It contains an attribute called ```key``` that stores a list of unique identifiers in the table. ```key``` attribute is similar to primary key attribute in SQL tables.  

### tuple

It is used to represent an entity in a table such as a product in walmart product dataset. It is represented as ```named list``` in R.

In QPackage a tuple can be created as follows

```{r}
t <- tuple("title" = "This is a title", "brand" = "Product brand")

# print the class
print(class(t))

# print contents of t
# title
print(t$title)
# brand
print(t$brand)
```

### Import data

As a first step, importing data from csv files are supported and it is done using ```read_csv``` command. 

The package is bundled with walmart and bowker book datasets as csv files. We can read them into qtables as follows

```{r}
# Get the path of bowker.csv file that is included along with the package
bowker_file_path <- system.file("extdata/books", "bowker.csv", package = "QPackage")
# Use read_csv to read the contents of csv file
bowker <- read_csv(bowker_file_path, status = read_status)

# print the status
print(read_status) #0 success, 1 error

# print the class
class(bowker)

```

Once the csv files are read into qtable, we can check if an attribute can be set as key attribute

```{r}
# Display colnames of bowker
colnames(bowker)
# check whether "id" can be set as key attribute
is_id <- check_id(bowker, list("id")) # 0 success, 1 error
# print the result
print(is_id)
```

Given that ```id``` can be set as key attribute, we can set it using ```set_id``` command

```{r}
# set "id" as key attribute
exit_status <- set_id(bowker, list("id"))
print(exit_status) # 0 success, 1 error
print(bowker@key)
```

#### Pre-loaded datasets

For convenience, QPackage is preloaded with bowker and walmart book datasets. The preloaded datasets can be attached to R session using the following command

```{r}
data(bowker_bk_dataset)
data(walmart_bk_dataset)
```
The above command loads bowker dataset to ```bowker_bk_dataset``` symbol and walmart dataset to ```walmart_bk_dataset```. The loaded datasets can be used as normal qtables.

```{r}
# assign to different variable
walmart <- walmart_bk_dataset
bowker <- bowker_bk_dataset

# print the class
class(bowker)

# print key attribute
print(bowker@key)
```
### Export data
As a first step, exporting data to csv file is supported and it is done using ```write_csv``` command.

```
status <- write_csv(bowker, 
          file="/scratch/pradap/downloads/bowker.csv") 
```

### View data

The loaded qtable can be viewed using the following command.

```
View(bowker) # output is suppressed
```

Currently it uses ```View``` command in R to display the contents.

### Blocking
Once the datasets are loaded, the user would like to do blocking as a step prior to matching. As a first step attribute equivalence based blocking is supported.

For instance, given that walmart and bowker datasets are loaded, the user may want to do blocking using ```isbn``` attribute and may want to include ```title```, ```author```, ```binding```, ```publisher``` and ```pages``` attributes in the candidate set. 

```{r}
# block attributes from walmart and bowker dataset 
attr1 <- "isbn" # from walmart dataset
attr2 <- "isbn" # from bowker dataset

# do blocking
cand_set <- apply_block(walmart, bowker, attr_equiv_block, attr1, attr2, 
                        col_names_a = list("title", "author", "binding", "publisher", "pages"), 
                        col_names_b = list("title", "author", "binding", "publisher", "pages")
                        )
# number of rows in candidate set
nrow(cand_set)

# attributes in candidate set
colnames(cand_set)
```

### Sample data

After blocking, the user may want to sample the candidate set to label them. As a first step random sampling is supported and it can done using the following command

```{r}
sample_data <- sample_qtable(cand_set, 25)
```

### Label data

As a next step, the user can label sampled data using GUI.  The sampled table is displayed to the user in a GUI  with an extra ```label``` column. 
The user can the ```label``` column to  mark a pair as match (```1```) or non-match (```0```).

```
labeled_cset <- label_data(sampled_data) # output is suppressed

``` 
Once the user finishes labelling, he can just close the browser (there is no explicit ```save``` command).

### Generating feature vectors

The labeled data can be used to generate feature vectors to train machine learning models. A feature is a similarity measure between a pair of attributes.

In QPackage a set of built-in similarity functions and tokenizers are supported and they can be viewed using the following commands

```{r}
show_builtin_simfuns()
show_builtin_tokenizers()
```

As a first step, given two input tables and their attribute correspondence, features are generated automatically using a set of heuristics based on attribute type and its properties, such as average number of words for character type. 
For instance, the user can generate features using ```title```, ```numAuthors``` and ```binding``` attributes from bowker and walmart table


```{r}
feature_list <- create_features(walmart, bowker, 
                                list(c("title", "title"), 
                                     c("numAuthors", "numAuthors"), 
                                     c("binding", "binding")))
length(feature_list)
```
Features are R functions that takes in two ```tuples``` and returns a numeric value

```{r}
print(feature_list[[2]])
```

```create_features``` can optionally take a list of allowed tokenizers and similarity functions as input, and restrict the features generated.

```{r}
test_list <- create_features(walmart, bowker, 
                                list(c("title", "title"), 
                                     c("numAuthors", "numAuthors"), 
                                     c("binding", "binding")), 
                                list("jaccard"), list("tok_qgram"))

length(test_list)
print(test_list)
```
Using the features, feature vector can be generated using the following command
```{r}
# ignore this command -- done only for illustration
data(test_label_cset) # this should not be executed.

labeled_feat_vec <- convert_to_feature_vecs(walmart, bowker, label_cset, feature_list)

# print column names
colnames(labeled_feat_vec)

# print first row
labeled_feat_vec[1, ]
```
### Model selection

Given labeled data and feature vectors, the user would like to choose a learning model for matching. He can choose a model by comparing accuracies from crossvalidation (CV).
For instance, the user decides to compare 3 learning models : (1) Decision tree, (2) Random forest, and (3) SVM. 

The accuracy can be computed using 10 fold CV as follows
```{r}
# Models compared : rpart, randomForest, SVM
# do 10 fold cross validation


# Assumption: .libPaths(c("/scratch/pradap/r-libs/", .libPaths())) 
# is already executed in the curret R - session
# install.packages("rpart")  ## required to install rpart
library(rpart)
acc_dt <- cv_kfold(labeled_feat_vec, 10, rpart, predict_args = list(type = "class"))

acc_dt

# Assumption: .libPaths(c("/scratch/pradap/r-libs/", .libPaths())) 
# is already executed in the curret R - session
#install.packages("randomForest") ## required to install randomFores
library(randomForest) 
acc_rf <- cv_kfold(labeled_feat_vec, 10, randomForest)

acc_rf

# Assumption: .libPaths(c("/scratch/pradap/r-libs/", .libPaths())) 
# is already executed in the curret R - session
# install.packages("e1071") ## required to install e1071
library(e1071)
acc_svm <- cv_kfold(labeled_feat_vec, 10, svm)

acc_svm

```

### Train model

Based on the accuracies, the user can select one of the models to train the whole labeled set. For instance, if the user chooses ```randomForest``` model then he can issue the following command to train it using whole labeled set

```{r}
model <- train_model(labeled_feat_vec, randomForest)

print(model)
```

### Predict labels

The trained model can be used to predict the matches in candidate set. 

The user can first generate feature vector for candidate set and can then use trained model to predict the matches.


```{r}
candset_feat_vec <- convert_to_feature_vecs(walmart, bowker, cand_set, feature_list)
candset_fv_with_labels<- predict_label(candset_feat_vec, model)

colnames(candset_fv_with_labels)
```

The candidate set with feature vectors has an extra column "predicted_labels" that contains predicted matches (```1```) and non-matches (```0```)

Optionally, the candidate set along with feature vectors can be viewed using the following command

````
View(merge(cand_set,candset_fv_with_labels)) # output is suppressed
````
